---
title: "R Notebook"
---

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidyr)
library(lvplot)
library(stringr)
library(kableExtra)
library(data.table)
library(sandwich)
library(lmtest)
library(jtools)
library(ggthemes)
library(huxtable)
library(RColorBrewer)
library(gt)
```

Import Data
```{r}
## Read the merged data of industry-level Execucomp and Annual Fundamentals

Industry_1 = read.csv("Final Matching Dataset.csv")
head(Industry_1)
Industry_1
```


**Summary Statistics for Outcome Variable -- Direction Executive**

Summary Table for Total/Current/Incentive Gender Compensation Gap
```{r}
## Total Compensation
Industry_1 %>%
  summarise(
    Min = min(gap_totcomp, na.rm = TRUE),
    Avg = mean(gap_totcomp, na.rm = TRUE),
    Max = max(gap_totcomp, na.rm = TRUE),
    `1st_Quantile` = quantile(gap_totcomp, 0.25, na.rm = TRUE),
    `3rd_Quantile` = quantile(gap_totcomp, 0.75, na.rm = TRUE),
    Std_Dev = sd(gap_totcomp, na.rm = TRUE)
  )%>%
  gt() %>%
  fmt_number(decimals = 2)

## Current Compensation
Industry_1 %>%
  summarise(
    Min = min(gap_currcomp, na.rm = TRUE),
    Avg = mean(gap_currcomp, na.rm = TRUE),
    Max = max(gap_currcomp, na.rm = TRUE),
    `1st_Quantile` = quantile(gap_currcomp, 0.25, na.rm = TRUE),
    `3rd_Quantile` = quantile(gap_currcomp, 0.75, na.rm = TRUE),
    Std_Dev = sd(gap_currcomp, na.rm = TRUE)
  )%>%
  gt() %>%
  fmt_number(decimals = 2)


## Incentive Compensation
Industry_1 %>%
  summarise(
    Min = min(gap_incencomp, na.rm = TRUE),
    Avg = mean(gap_incencomp, na.rm = TRUE),
    Max = max(gap_incencomp, na.rm = TRUE),
    `1st_Quantile` = quantile(gap_incencomp, 0.25, na.rm = TRUE),
    `3rd_Quantile` = quantile(gap_incencomp, 0.75, na.rm = TRUE),
    Std_Dev = sd(gap_incencomp, na.rm = TRUE)
  )%>%
  gt() %>%
  fmt_number(decimals = 2)
```

Histogram of Outcome Variable "Total Compensation Gap"
```{r}
## To better display, only show distribution betwee 0.01-0.99 quantile
lower <- quantile(Industry_1$gap_totcomp, 0.01, na.rm=T)
upper <- quantile(Industry_1$gap_totcomp, 0.99, na.rm=T)

Industry_1 %>%
  ggplot(aes(x=gap_totcomp)) +
  geom_histogram(fill=(brewer.pal(9, "Blues")[9]), color="black", bins=50) +
  labs(title="Distribution of Average Total Industry Compensation Gap",
       x="Total Compensation Gap",
       y="Frequency") +
  coord_cartesian(xlim = c(lower, upper)) +  
  theme_stata(scheme = "s1color") +
  scale_colour_stata(scheme = "s2color")+
  theme(
        title = element_text(hjust = 0.5,face="bold")
        )
```

Trend of three compensation type's gap
```{r}
# Show the compensatioon gap over years for total compensation, current compensation, and incentive compensation

Industry_1 %>%
  group_by(year) %>%
  summarize(
    average_totalgap = mean(gap_totcomp, na.rm=T),
    average_currgap = mean(gap_currcomp, na.rm=T),
    average_incengap = mean(gap_incencomp, na.rm=T)
  ) %>%
  ggplot(aes(x=year)) +
  geom_line(aes(y = average_totalgap, color = "Total Gap"), size = 1) +  
  geom_line(aes(y = average_currgap, color = "Current Gap"), size = 1) +
  geom_line(aes(y = average_incengap, color = "Incentive Gap"), size = 1) +
  labs(
    title = "Average Compensation Gap Trend: 2012-2022",
    x = "Year",
    y = "Average Compensation Gap between Genders",
    colour = ""
  ) +
  theme_stata(scheme = "s1color") +
  scale_colour_stata(scheme = "s2color") + 
  scale_shape_stata() +
  theme(axis.title = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        title = element_text(hjust = 0.5, face="bold"))+
        scale_x_continuous(breaks = seq(min(Industry_1$year), max(Industry_1$year), by = 1))
```

Summary Statistics for Executive Compensation Gap by Industry & Sector
```{r}
Industry_1 %>%
  group_by(ind2dig) %>%
  summarize(
    mean_totgap = mean(gap_totcomp, na.rm=T), 
    mean_currgap = mean(gap_currcomp, na.rm=T),
    mean_incengap = mean(gap_incencomp, na.rm=T)
  ) %>%
  gt()%>%
  fmt_number(decimals = 2)

Industry_1 %>%
  group_by(sector) %>%
  summarize(
    mean_totgap = mean(gap_totcomp, na.rm=T), 
    mean_currgap = mean(gap_currcomp, na.rm=T),
    mean_incengap = mean(gap_incencomp, na.rm=T)
  ) %>%
  gt()%>%
  fmt_number(decimals = 2)
```

**Summary Statistics for Outcome Variable -- Gender Wage Gap**
```{r}
## Gender Wage Gap (ln)
Industry_1 %>%
  summarise(
    Min = min(lgapearn, na.rm = TRUE),
    Avg = mean(lgapearn, na.rm = TRUE),
    Max = max(lgapearn, na.rm = TRUE),
    `1st_Quantile` = quantile(lgapearn, 0.25, na.rm = TRUE),
    `3rd_Quantile` = quantile(lgapearn, 0.75, na.rm = TRUE),
    Std_Dev = sd(lgapearn, na.rm = TRUE)
  )%>%
  gt() %>%
  fmt_number(decimals = 2)
```


Histogram of Outcome Variable- non-managerial labor wage gap (ln)
```{r}
## To better display, only show distribution betwee 0.01-0.99 quantile
lower_earn <- quantile(Industry_1$lgapearn, 0.01, na.rm=T)
upper_earn <- quantile(Industry_1$lgapearn, 0.99, na.rm=T)

ggplot(Industry_1, aes(x=lgapearn)) +
  geom_histogram(fill=brewer.pal(9, "Blues")[9], color="black", bins=50) +
  labs(title="Distribution of Industry Gender Wage Gap",
       x="Log of Wage Gap",
       y="Frequency") +
  coord_cartesian(xlim = c(lower_earn, upper_earn)) +
  theme_stata(scheme = "s1color") +
  scale_colour_stata(scheme = "s2color")+
  theme(
        title = element_text(hjust = 0.5,face="bold")
        )
```

Summary Statistics for Gender Wage Gap by Industry & Sector
```{r}
Industry_1 %>%
  group_by(ind2dig) %>%
  summarize(
    earn_gap = mean(lgapearn, na.rm=T)
  ) %>%
  gt()%>%
  fmt_number(decimals = 2)

earn_gap = Industry_1 %>%
  group_by(sector) %>%
  summarize(
    earn_gap = mean(lgapearn, na.rm=T)
  )


ggplot(earn_gap, aes(x=reorder(sector, -earn_gap), y=earn_gap)) +
  geom_bar(stat="identity", fill="steelblue") +
  coord_flip() +
  labs(title="Distribution of Industry Gender Wage Gap",
       x="Sector",
       y="Log of Wage Gap") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"),
        axis.text.x = element_text(size=10),  # Adjust text size as needed
        axis.text.y = element_text(size=10))
```


**Summary Statistics for Treatment Variable -- Female CEO Share**
Summary Statistics for male and female share:
```{r}
## Summary of female CEO share by sector
Industry_1 %>%
  group_by(sector) %>%
  summarise(
    Min = min(share_ceo, na.rm = TRUE),
    Avg = mean(share_ceo, na.rm = TRUE),
    Max = max(share_ceo, na.rm = TRUE),
    `1st_Quantile` = quantile(share_ceo, 0.25, na.rm = TRUE),
    `3rd_Quantile` = quantile(share_ceo, 0.75, na.rm = TRUE),
    Std_Dev = sd(share_ceo, na.rm = TRUE),
    Num_NA = sum(is.na(share_ceo))
  ) %>%
  gt() %>%
  fmt_number(decimals = 4)

## Boxplot visualization
CEOshare = Industry_1 %>%
  group_by(sector) %>%
  summarise(
    Min = min(share_ceo, na.rm = TRUE),
    Avg = mean(share_ceo, na.rm = TRUE),
    Max = max(share_ceo, na.rm = TRUE),
    first_Quantile = quantile(share_ceo, 0.25, na.rm = TRUE),
    thrid_Quantile = quantile(share_ceo, 0.75, na.rm = TRUE),
    Std_Dev = sd(share_ceo, na.rm = TRUE),
    Num_NA = sum(is.na(share_ceo))
  ) 

ggplot(CEOshare, aes(x=reorder(sector, -Avg), y=Avg)) +
  geom_boxplot(aes(lower=first_Quantile, upper=thrid_Quantile, middle=Avg, ymin=Min, ymax=Max), stat="identity", fill="lightblue") +
  coord_flip() +
  labs(title="Summary Statistics of Industry Female CEO Share",
       x="Industry Sector",
       y="Female CEO Share") +
  theme_minimal() +
  theme(plot.title = element_text(face="bold", hjust = 0.5),
        axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(face="bold"),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

Distribution plot (Histogram) of female executive share
```{r}
Industry_1 %>%
  ggplot(aes(x=share_ceo)) +
  geom_histogram(fill=(brewer.pal(9, "Blues")[9]), color="black", bins=50) +
  labs(title="Distribution of Industry's Female CEO Share",
       x="Female CEO Share",
       y="Frequency") +
  theme_stata(scheme = "s1color") +
  scale_colour_stata(scheme = "s2color")+
  theme(
        title = element_text(hjust = 0.5,face="bold")
        )
```
